Untitled

### 2020-05-11 tour-vis.Rmd
more vis method and development

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Sherry Zhang 
    url: https://github.com/huizezhang-sherry
    affiliation: Monash University
    affiliation_url: https://www.monash.edu/
date: "`r Sys.Date()`"
output: distill::distill_article
bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
load(here::here("data/temp.rda"))
```

@@

Simulated annealing @lee2005projection @klein1989experiments is a non-derivative procedure based on a cooling *temperature* parametere $T_i$. Given an initial $T_0$, the temperature at iteration $i$ is defined as $T_i = \frac{T_0}{log(i + 1)}$. The simulated annealing algorithm works as follows. Given a proportion $p$ and a randomly generated orthonormal basis $B'$, a candidate basis is constructed as $(1- p)B + p B'$ where $B$ is the current basis. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(\frac{I(B) - I(B')}{T_i}), 1 \right)$ where $I(.)$ is the index function.

    [The mathmetical expression is from Lee 2005 - without a negative in the $\exp$ but in the initial paper where SA is proposed 
    @article{bertsimas1993simulated,
      title={Simulated annealing},
      author={Bertsimas, Dimitris and Tsitsiklis, John and others},
      journal={Statistical science},
      volume={8},
      number={1},
      pages={10--15},
      year={1993},
      publisher={Institute of Mathematical Statistics}
    }
    ,

    @article{kirkpatrick1983optimization,
      title={Optimization by simulated annealing},
      author={Kirkpatrick, Scott and Gelatt, C Daniel and Vecchi, Mario P},
      journal={science},
      volume={220},
      number={4598},
      pages={671--680},
      year={1983},
      publisher={American association for the advancement of science}
    }

    and

    @article{klein1989experiments,
      title={Experiments in projection and clustering by simulated annealing},
      author={Klein, Raymond W and Dubes, Richard C},
      journal={Pattern Recognition},
      volume={22},
      number={2},
      pages={213--220},
      year={1989},
      publisher={Elsevier}
    }

    The definition of $T_i$ is $T_i = \exp(-\frac{\Delta I}{k T})$ - with negative in the $\exp$. It is the gibb/boltzman distribution!]

@@

The global structure consists of six columnes: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process. The first five rows of the record of a five-variable 1d projection is printed below.

```{r eval = FALSE}
devtools::load_all()
set.seed(123456)
temp <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_geodesic),
                  rescale = FALSE, verbose = TRUE)

@@

`tries` has an increment of one once the generator is called (equivalently a new target basis is generated); `info` records the stage the basis is in. This would include the `interpolation` stage and the detailed stage in the optimisation i.e. `direction_search`, `best_direction_search`, `line_search`and `best_line_search` for geodesic searching (`search_geodesic`); `random_search` and `new_basis` for simulating annealing (`search_better`). `loop` is the counter used for the optimisation procedure and thus will be `NA` for interpolation steps. `id` creates a sequential order of the basis. This information will be stored and printed when the optimisation ends and can be turned off via `print = FALSE`. Additional messages during the optimisation can be displayed via `verbose = TRUE`.

# Visual methods
## index vs time plots
## target basis vs interpolation steps, index vs time
## low-d representation, PCA

@@

# Finding errors and developing improvements
## interrupt
## polish

>>>>

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Sherry Zhang 
    url: https://github.com/huizezhang-sherry
    affiliation: Monash University
    affiliation_url: https://www.monash.edu/
date: "`r Sys.Date()`"
output: distill::distill_article
bibliography: biblio.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ferrn)
library(patchwork)

load(here::here("data/temp.rda"))
load(here::here("data/data_mult.rda"))
load(here::here("data/holes_1d_geo.rda"))
load(here::here("data/holes_2d_geo.rda"))
load(here::here("data/holes_2d_better.rda"))

load(here::here("data/holes_2d_geo_polish.rda"))
load(here::here("data/holes_2d_better_polish.rda"))

load(here::here("data/interrupt_yes.rda"))
load(here::here("data/interrupt_no.rda"))
```

@@

Simulated annealing @bertsimas1993simulated @kirkpatrick1983optimization is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$. Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.

@@

Two set of simulated data are used in the demonstration of the visualiation and diagnostics of the tour optimisation. A small dataset consists of 1000 randomly simulated observations of five variables (`x1`, `x2`, `x8`, `x9`, `x10`). `x2` is the informative variable simulated from two bi-modal normal distribution centered at -3 and 3 with variance being 1 and the other four are simulated from `N(0, 1)`. The data has been scaled to ensure `x2` has variance of 1.

A larger dataset contains more informative variables (`x3` to `x7`) of different types. `x3` takes 500 positive one and 500 negative one. The distribution of all the variables except `x3` is plotted below. *[should I introduce the dist for each var?]*

```{r}
origin_dt_bi <- data_mult %>% 
  dplyr::select(-x3) %>% 
  gather(names, values) %>%
  mutate(names = as_factor(names))

origin_dt_bi %>%
  ggplot(aes(x = values)) +
  geom_histogram(binwidth = 0.3) +
  geom_density(aes(y = 0.3 * ..count..)) +
  facet_wrap(vars(names), ncol = 3)
```

Once the dataset is sent to the tourr package, all the information generated will be stored in a global structure. The global structure consists of six columnes: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process. The example below presents the global object of a 1D projection of the small dataset with geodesic seraching method.

```{r eval = FALSE}
set.seed(123456)
temp <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_geodesic),
                  rescale = FALSE, verbose = TRUE)

@@

`tries` has an increment of one once the generator is called (equivalently a new target basis is generated); `info` records the stage the basis is in. This would include the `interpolation` stage and the detailed stage in the optimisation i.e. `direction_search`, `best_direction_search`, `line_search`and `best_line_search` for geodesic searching (`search_geodesic`); `random_search` and `new_basis` for simulating annealing (`search_better`). `loop` is the counter used for the optimisation procedure and thus will be `NA` for interpolation steps. `id` creates a sequential order of the basis. This information will be stored and printed when the optimisation ends and can be turned off via `print = FALSE`. Additional messages during the optimisation can be displayed via `verbose = TRUE`. Another examples is a 2D projection of the larger dataset with two informative variable (`x2` and `x7`) using search_better method. Notice in this example, the dimension of the bases becomes 6 by 2.

```{r echo = TRUE}
holes_2d_better %>% head(5)
```

# Visual methods

The creation of the global object facilitates the diagnostics of the optimisation in the projection pursuit guided tourr. Different diagnostic plots could be made to understand the optimisation **[pretty cheesy here]**

*The examples and results below still need to be modified as the development / modification of both ferrn and tourr package*.

## index vs time plots

```{r echo = TRUE}
holes_2d_geo %>% explore_trace_all()
```

## target basis vs interpolation steps, index vs time

```{r echo = TRUE}
holes_2d_geo %>% explore_trace_interp()
holes_2d_better %>% explore_trace_interp()

interrupt_yes %>% explore_trace_interp()
```

## low-d representation, PCA

```{r echo = TRUE}
holes_1d_geo %>% explore_proj_pca()
```

@@

```{r echo = TRUE}
holes_1d_geo %>% explore_proj_pca(animate = TRUE)
```

# Finding errors and developing improvements

## interrupt

```{r}
p1 <- interrupt_no %>% explore_trace_interp() + ggtitle("without interruption")

p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption")

p1 | p2
```

Further the case when the index function is not smooth, the interpolation may not attain -> modification

## Polish

In principle, all the optimisation routines should present the same output on the same problem. However, as we observed, the results from using `search_better` and `search_geodesic` on our 2D projection problem are different. Although both methods have detected `x2` as an informative variable on the second column, the weights for each variable on the first column vary, which causes difference on the index value. This motivates us to create another searching function that polishes the ending basis of the previous search and achieve unity on different methods.

```{r echo = TRUE}
holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]

holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(index_val) 
holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(index_val) 
```

**[Talk about how search_polish works]**

The usage of search_polish is as follows. After the first tour, the final basis from the interpolation is extracted and supplied intoa new tour with the `start` argument and `search_polish` as teh searching function in the guided_tour. All the other arguments should remain the same.

```{r eval = FALSE, echo = TRUE}
set.seed(123456)
holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)],
                           tour_path = guided_tour(holes(), d = 2, 
                                                   search_f = tourr:::search_geodesic),
                           rescale = FALSE, verbose = TRUE)

last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]

set.seed(123456)
holes_2d_geo_polish <- animate_xy(data_mult[,c(1,2, 7:10)],
                                     tour_path = guided_tour(holes(), d = 2, 
                                                             search_f = tourr:::search_polish),
                                     rescale = FALSE, verbose = TRUE, start = last_basis)
```

**[This example is not ideal because currently the geodesic is not doing a good job in the first place]**

```{r echo = TRUE}
holes_2d_better_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(index_val)
holes_2d_geo_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(index_val)

holes_2d_better_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(basis) %>% .[[1]]
holes_2d_geo_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(basis) %>% .[[1]]

```

## posse?

search_better annd search_geodesic methods are not doing good on 2d projection -> see the simple case above: abs best is 0.96 while they only get to 0.88-ish. Would be worse for higher dimension. thus search_posse. get 0.95 and then polish to 0.96 :)
